### 2023-10-31 cluster_analysis.qmd
first commit

---
title: "Formulating Regional Typologies"
author: "Dony Indiarto, Mohamad Nugraha"
affiliation: "CIFOR-ICRAF Indonesia"
format: revealjs
editor: visual
---

## Agenda

1. Understanding the Regional Typologies
2. Methodology and Steps
3. Case Studies and Examples
4. Closing Remarks

------------------------------------------------------------------------

## What are Regional Typologies?

-   **Definition**
    -   A categorization of geographical areas based on shared features.
    -   Features often include quantifiable socio-economic indicators and environmental factors.

------------------------------------------------------------------------

## Why are Regional Typologies Important?

- Regions vary in socio-economic and environmental aspects.
- Typologies simplify this by creating manageable groups.
- They facilitate comparisons and targeted management.

------------------------------------------------------------------------
## How Do Typology and Regional Indexes Differ? {.smaller}

:::: {.columns}

::: {.column width="50%"}
### Urban-rural typology
<!-- ![Urban-rural typology for NUTS level 3 regions](image/nuts.png) -->
:::
### Regional Index
::: {.column width="50%"}
<!-- ![Global Food Security Index](image/gfsi.png) -->
:::

::::

------------------------------------------------------------------------

## How Do Typology and Regional Indexes Differ? {.smaller}

| Criteria                  | Regional Typologies                                                                                 | Regional Index                                                                               |
|----------------|--------------------------------|-------------------------|
| **What it Is**            | Categorises areas based on similar socio-economic, environmental, or institutional characteristics. | A single, composite measure that quantifies the level of vulnerability to a specific threat. |
| **Outcome**               | Classifications like "high-income urban area" or "low-resource rural area."                         | Numerical score ranking areas from low to high vulnerability.                                |
| **Use-Case**              | Tailoring policies or interventions to different types of areas.                                    | Prioritising aid, resources, or interventions based on vulnerability.                        |
| **Caveats**               | Can oversimplify complexities; subject to analyst bias.                                             | May not capture the full situation; choice of indicators and weightings can be arbitrary.    |
| **Complementary Aspects** | Provides rich, contextual information.                                                              | Pinpoints areas needing immediate attention.                                                 |

------------------------------------------------------------------------

## How Do We Make Regional Typologies?

-   Formulate Objectives
-   Identify  Data requirement and availability
-   Data pre-processing
-   Variable selection and PCA
-   Cluster analysis and validation
-   Interpretation and comparisons

------------------------------------------------------------------------

## Slide 6: Case Studies and Examples

-   **Previous Projects**
    -   West Kalimantan
    -   South Sumatra
-   **Focus: South Sulawesi**
    -   Scope, Objectives, and Key Findings

------------------------------------------------------------------------

## Slide 7: Summary & Open Question

-   **Summary**
    -   Importance of Cluster Analysis in regional planning.
-   **Open Question**
    -   What would you like to learn from a cluster analysis for your specific context?

------------------------------------------------------------------------