### 2023-11-01 cluster_analysis.qmd
updated slides

---
title: "Formulating Regional Typologies"
author: "Dony Indiarto, Mohamad Nugraha"
affiliation: "CIFOR-ICRAF Indonesia"
project:
  type: website
  output-dir: docs
format: 
  revealjs:
    theme: simple
    scrollable: true
    embed-resources: true
reference-location: document
editor: visual
---

## Agenda

1.  Understanding the Regional Typologies
2.  Methodology and Steps
3.  Case Studies and Examples
4.  Closing Remarks

------------------------------------------------------------------------

## What are Regional Typologies?

-   **Definition**
    -   A categorization of geographical areas based on shared features.
    -   Features often include quantifiable socio-economic indicators and environmental factors.

------------------------------------------------------------------------

## Why are Regional Typologies Important?

-   Regions vary in socio-economic and environmental aspects.
-   Typologies simplify this by creating manageable groups.
-   They facilitate comparisons and targeted management.

------------------------------------------------------------------------

## How Do Typologies and Indexes Differ? {.smaller}

::: columns
::: {.column width="50%"}
**Typology**

![Urban-rural typology for NUTS level 3 regions](image/nuts.png){height="400"}
:::

::: {.column width="50%"}
**Index**

![Global Food Security Index](image/gfsi.png){height="400"}
:::
:::

------------------------------------------------------------------------

## How Do Typologies and Indexes Differ? {.smaller}

::: {style="font-size:20px"}
| Criteria                  | Regional Typologies                                                                                 | Regional Indexes                                                                          |
|------------------|------------------------------|------------------------|
| **What it is**            | Categorises areas based on similar socio-economic, environmental, or institutional characteristics. | A single, composite measure that quantifies the level of vulnerability to a threat.       |
| **Outcome**               | Classifications like "urban area" or "rural area."                                                  | Numerical score ranking areas from low to high vulnerability.                             |
| **Use-case**              | Tailoring policies or interventions to different types of areas.                                    | Prioritising aid, resources, or interventions based on the ranking                        |
| **Caveats**               | Can oversimplify complexities; subject to analyst bias.                                             | May not capture the full situation; choice of indicators and weightings can be arbitrary. |
| **Complementary aspects** | Provides rich, contextual information.                                                              | Pinpoints areas needing immediate attention.                                              |

: {.striped .hover}
:::

------------------------------------------------------------------------

## How Do We Make Regional Typologies?

::: {.panel-tabset style="font-size:20px"}
### Key steps

-   Formulate Objectives
-   Identify Data requirement and availability
-   Data collection & pre-processing
-   Variable selection and PCA[^1]
-   Cluster analysis[^2] and validation
-   Interpretation and comparisons

### Technical Workflow

![](image/workflow.svg){height="1200" fig-align="center"}
:::

[^1]: Principle Component Analysis is a statistical technique used to simplify the complexity in data by highlighting its most important features.

[^2]: Cluster analysis is a machine learning procedure used to group similar items together based on their characteristics.

------------------------------------------------------------------------

# Case Studies

------------------------------------------------------------------------

## Case Study #1 {.smaller}

::: {style="font-size:18px"}
**Assessing and Characterizing Climate Vulnerabilities in Agriculture-Based Livelihoods: South Sulawesi**
:::

::: {.panel-tabset style="font-size:16px"}
### Description

**Aim**: Analyze vulnerabilities of agriculture livelihoods at the provincial level; identify key risks, root causes, and adaptation strategies.

**Approach**: Define locations with similar biophysical and socio-economic characteristics as 'typologies' using PCA & K-means clustering.

**Area of Interest**: South Sulawesi province

**Unit of Analysis**: Sub-district (Kecamatan)

### List of potential features

### Collected data

### PCA Summary

### K-means cluster
:::

------------------------------------------------------------------------

## Case Study #1 {.smaller}

::: {style="font-size:18px"}
**Assessing and Characterizing Climate Vulnerabilities in Agriculture-Based Livelihoods: South Sulawesi**
:::

::: {.panel-tabset style="font-size:16px"}
### Map

### **1. Islands**

-   The second lowest household density after rural areas.

-   Limited accessibility to land-based livelihoods.

-   Economy predominantly relies on marine resources and maritime transport.

-   Inadequate health, education facilities, and market access.

-   Driest area in South Sulawesi with potential freshwater scarcity.

-   High drought indices with significant predicted rainfall decrease.

-   Over 40% of the population is underprivileged.

-   Mostly situated in warm lowlands; a 1°C temperature rise could be impactful.

### **2. Predominantly Rural & Biodiversity Rich**

-   Rural areas in South Sulawesi with the lowest household density.

-   Mountainous regions overlapping with protected areas, last bastions of unique Wallacean biodiversity.

-   High rainfall, low drought vulnerability (AI = 0.3).

-   In close proximity to forests, deforestation sites, and areas previously affected by fires.

-   Furthest from plantation concessions and limited road and river access.

-   Minimal changes in rainfall but the highest rate of temperature increase.

-   Highest erosion risk; lowest flood risk.

-   Almost 50% of the population falls in the bottom 40% economic bracket.

### **3. Predominantly Plantation-Driven Economy**

-   Closest to plantations, forests, and deforestation sites.

-   Furthest from previously burned areas in Sulawesi.

-   Medium household density per sub-district.

-   Moderate flood risk.

-   Moderate extent of irrigation.

-   Second highest risk of erosion and landslides after inland areas.

-   Highest rate of deforestation.

-   High risk of rainfall reduction, though not classified as arid.

-   About 1/3 of the population falls in the bottom 40% economic bracket.

### **4. Food Basket**

-   Highest arable land coverage, with predominant irrigated areas.

-   Highest smallholder land ownership.

-   Second closest proximity to burned areas after urban regions.

-   Second highest access to roads and rivers following urban areas.

-   Second highest household density per district after urban regions; similar trend observed in health and education facilities.

-   Least proximity to forests due to limited forest cover.

-   Primarily located in lowlands; a temperature rise of 1.41°C significantly impacts this already warm region.

-   Expected to experience the most significant rainfall decrease (-84.31 mm) among typologies, making it the most drought-prone.

-   Flood risks during the rainy season, aggravated by minimal remaining forest cover.

-   High household density, with over a third belonging to the bottom 40% economic bracket.

### **5. Predominantly Industry & Service-Driven Economy**

-   Smallest area coverage (3%) with least arable land.

-   Lowest smallholder land ownership, following island regions.

-   Second closest proximity to burned areas after semi-urban regions.

-   Best access to roads and rivers.

-   Highest household density per district.

-   Easy access to health facilities, supermarkets, universities, and hospitals. However, each unit caters to a large number of households.

-   Only 15.1% of households fall within the bottom 40% economic bracket.

-   Warmest current temperatures compared to other regions; a rise of 1.43°C significantly impacts underprivileged local residents.

-   High flood risk, given many urban areas are near rivers.

-   Heavily reliant on food supplies from other regions.

### Area Composition

### Demographic Composition
:::

------------------------------------------------------------------------

## Case Study #2 {.smaller}

::: {style="font-size:18px"}
**Village-level vulnerability classes of agricultural-based livelihoods to climate change in the studied districts in West Kalimantan Province**
:::

::: {.panel-tabset style="font-size:16px"}
### Map

![](image/typ_west_kal.PNG){height="450" fig-align="center"}

### Description

**1 -- Most Vulnerable**: Located closest to oil palm plantations and factories, mining areas, and roads; has the largest shrub area per village; furthest from deforestation and has the lowest deforestation rate; has the smallest percentage of forested areas per village; high population.

**2 -- Highly Vulnerable**: Located closest to burnt areas; has the largest percentage of oil palm area per village; situated nearer to oil palm concession areas and rubber factories; has the largest percentage of water bodies (lakes, rivers); has a lower deforestation rate but is located closer to deforestation areas; high population.

**3 -- Moderately Vulnerable**: Located closest to rivers; has a larger percentage of oil palm area per village; situated slightly farther from oil palm companies and mining areas; has a larger percentage of forested and shrub areas per village; medium population.

**4 -- Less Vulnerable**: Closest to deforestation areas and has the highest deforestation rate; located furthest from rivers; most distant from burnt areas; slightly closer to forested areas; slightly farther from oil palm concession areas; low population.

**5 -- Least Vulnerable**: Has the largest percentage of forested areas per village; most remote; has the smallest percentage of shrub areas and oil palm areas per village; located closer to rivers and forested areas; has the lowest village population.

### Source of Data

| No  | Category                      | Spatially explicit variables                              | References                                                                   |
|----------------|----------------|-----------------|------------------------|
| 1   | Distance to infrastructure    | Distance to oil palm plantation                           | Landcover map 2017 from MoEF                                                 |
| 2   | Distance to infrastructure    | Distance to roads                                         | BIG                                                                          |
| 3   | Distance to infrastructure    | Distance to a rubber factory                              | ICRAF                                                                        |
| 4   | Distance to infrastructure    | Distance to oil palm factory                              | ICRAF                                                                        |
| 5   | Distance to natural resources | Distance to forest                                        | Landcover map 2017 from MoEF                                                 |
| 6   | Distance to natural resources | Distance to river                                         | BIG                                                                          |
| 7   | Distance to natural resources | Distance to mining areas                                  | BIG                                                                          |
| 8   | Distance to hazards area      | Distance to burnt areas                                   | KLHK 2015                                                                    |
| 9   | Land use and land cover       | \% area of oil palm per village                           | Landcover map 2017 from MoEF                                                 |
| 10  | Land use and land cover       | \% of forested areas per village                          | Landcover map 2017 from MoEF                                                 |
| 11  | Land use and land cover       | \% of shrubs areas per village                            | Landcover map 2017 from MoEF                                                 |
| 12  | Land use and land cover       | \% forested areas per district compared to district areas | Landcover map 2017 from MoEF                                                 |
| 13  | Land use and land cover       | \% oil palm areas at district level compare               | Landcover map 2017 from MoEF                                                 |
| 14  | Land use and land cover       | \% water body compared to district areas                  | Badan Informasi Geospatial (BIG), Geospatial Information Agency of Indonesia |
| 15  | Land use and land cover       | Distance to deforestation                                 | Landcover map 2017 from MoEF                                                 |
| 16  | Land use and land cover       | Deforestation area                                        | Landcover map 2017 from MoEF                                                 |
| 17  | Hazards                       | Flood incidence                                           | Village Potentials, Statistics Indonesia 2018                                |
| 18  | Hazards                       | Heavy flood incidence                                     | Village Potentials, Statistics Indonesia 2018                                |
| 19  | Hazards                       | Fire incidences                                           | Village Potentials, Statistics Indonesia 2018                                |
| 20  | Hazards                       | Drought incidences                                        | Village Potentials, Statistics Indonesia 2018                                |
| 21  | Demography                    | Village population                                        | Village Potentials, Statistics Indonesia 2018                                |
| 22  | Distance to infrastructure    | Distance to oil palm concession areas                     | Department of Estate Crops, West Kalimantan Province                         |

: {.striped .hover}
:::

------------------------------------------------------------------------

## Case Study #3 {.smaller}

::: {style="font-size:18px"}
**Peat Ecosystem Typologies in South Sumatra Province**
:::

::: {.panel-tabset style="font-size:16px"}
### Map

![](image/typ_south_sum.PNG){height="450" fig-align="center"}

### Description

**Typology 1**: High fire occurrences in protected areas, both with and without drainage. Lowest rates of deforestation, degradation, and plantation expansion. Low population and poverty, closest to cities with easy access.

**Typology 2**: High fire occurrences in cultivated areas, both with and without drainage. High deforestation rate, low degradation, and minimal plantation expansion. Highest population and poverty, near cities with easy access.

**Typology 3**: Fires occur in both cultivated and protected areas, irrespective of drainage. Highest rates of deforestation, degradation, and plantation expansion. Low population and poverty, furthest from cities with challenging access.

**Typology 4**: Most frequent fires in protected areas, regardless of drainage. Low deforestation, high degradation, and significant plantation expansion. Low population and poverty, far from cities with very limited access.

**Typology 5**: Most frequent fires in cultivated areas, irrespective of drainage. Low deforestation, high degradation, and significant plantation expansion. Low population and poverty, remote from cities with challenging access.

### Typology-Based Intervention

| **Typology** | **R.3.1 Establishment of Peatland Stewardship Village**                                                                                                                                                                                                               | **R.3.2 Enhancing Livelihood Capacity**                                                                                                          | **R.3.3 Development of Alternative Commodities and Livelihood Sources**                                                             |
|----------------|----------------|----------------|------------------------|
| **1**        | \- Community awareness about the importance of the peat ecosystem, collaborating with district and provincial governments. <br><br> - Activities of the Peatland Stewardship Village aimed at increasing income and creating sources of livelihood for the community. | Conducted in synergy with district/city governments.                                                                                             | Directed at non-agricultural livelihood sources (off-farm). Local industries that exist need support.                               |
| **2**        | Peatland Stewardship Village is executed in an integrated manner in partnership with HTI companies and palm oil plantations. The goal is to enhance community access.                                                                                                 | Strengthening farmers' livelihoods to protect the remaining peat ecosystems. Formulation of alternative livelihood sources.                      | Livelihood sources oriented towards peat-friendly agriculture, mixed plantations combined with non-agricultural livelihood sources. |
| **3**        | Peatland Stewardship Village is oriented towards fire prevention; conducted in an integrated manner in partnership with HTI companies and palm oil plantations.                                                                                                       | Strengthening farmers' livelihoods through partnerships with plantation and HTI companies. Livelihoods aimed at forest and land fire prevention. | Livelihood sources oriented towards peat-friendly agriculture and plantations combined with fisheries and livestock.                |
| **4**        | Peatland Stewardship Village geared towards fire prevention; enhancing community access to markets and high-economic value commodities.                                                                                                                               | Strengthening farmers' livelihoods through partnerships with plantation and HTI companies. Emphasis on preventing forest and land fires.         | Livelihood sources focused on peat-friendly agriculture and plantations combined with fisheries and livestock.                      |
| **5**        | Peatland Stewardship Village focused on fire prevention; enhancing community access to markets and commodities of high economic value.                                                                                                                                | Strengthening farmers' livelihoods through partnerships with plantation and HTI companies. Emphasis on preventing forest and land fires.         | Livelihood sources focused on peat-friendly agriculture and plantations combined with fisheries and livestock.                      |

: Classification of peat ecosystem revitalization actions in South Sumatra province {.striped .hover}
:::

------------------------------------------------------------------------

## Conclusion

-   We showcase examples on how regional typologies can be produced
-   The aim of the study shapes the data selection and interpretation of the typologies.
-   Iterative feedback from experts is essential for ensuring typologies align with objectives.

------------------------------------------------------------------------